---
title: "How well is the exercise?"
author: "Widya Kartika"
date: "March 28, 2018"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1 Excecutive Summary

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.The goal of this assignment is to predict the manner in which they did the exercise. Using randomforest algorithm we create a prediction model with accuracy:

### 2 Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, our goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different way, thus 5 diferent classe variables.

More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

### 3 Getting and Cleaning Data

#### 3.1 Load Data

We load the training and test data with defining that na.string contains: (1)empty string (""), (2) "NA" and (3) "#DIV/0!".

```{r loadData}

pml.training <- read.csv("C:/Users/widya/Desktop/Coursera/Machine Learning/project/pml-training.csv",
                         header=T, na.strings=c("","#DIV/0!","NA"))
pml.testing <- read.csv("C:/Users/widya/Desktop/Coursera/Machine Learning/project/pml-testing.csv",
                         header=T, na.strings=c("","#DIV/0!","NA"))
```


#### 3.2 Clean Data

The column with NA is identified as follows:
```{r check_NA}
pml.training.na <- apply(pml.training,2,function(f) sum(is.na(f)))
pml.testing.na <- apply(pml.testing,2,function(f) sum(is.na(f)))
length(pml.training.na[ pml.training.na > 1])
min(pml.training.na[ pml.training.na > 1])
```

Then we remove all those 100 columns with NA > 19216 20 for testing
```{r remove_NA}
pml.training <-  pml.training[,pml.training.na < 19216] 
pml.testing <-  pml.testing[,pml.testing.na < 20] 
```

We also remove irrelevant column 1-7
```{r remove_irrelevant_coulumns}
pml.training[1:2,c(1:7)]
pml.training <-  pml.training[,-c(1:7)] 
pml.testing <-  pml.testing[,-c(1:7)] 
dim(pml.training)
dim(pml.testing)
```



### 4 K-fold cross validation

#### 4.1  Load Library and Setting
```{r libs, results='hide'}
set.seed(0)
library(caret)
library(randomForest)
library(party)
```

#### 4.2 Perform 10-fold cross validation as follows:

```{r 10-folds}
k <- 10
y <- c(1:nrow(pml.training))
folds <- createFolds(y, k, list = TRUE, returnTrain = FALSE)
obs <- list()
pred <- list()
for(i in 1:k){
  Indexes <- folds[[i]]
  testData <- pml.training[Indexes, ]
  trainData <- pml.training[-Indexes, ]
  #Use the test and train data partitions
  rf <- randomForest(classe ~ ., data=trainData)
  obs[[i]] <- testData$classe
  pred[[i]] <- predict(rf, testData)
}

```


#### 4.3 Result

```{r result}
accuracy = c()
for(i in 1:k){
  conf_mat <- confusionMatrix(pred[[i]], obs[[i]])
  accuracy <- c(accuracy,conf_mat$overall[1])
}

```

The accuracy range
Max:
```{r range_max, echo=FALSE}
max(accuracy)
```
Min:
```{r range, echo=FALSE}
min(accuracy)
```

```{r plot_range, echo=FALSE}
plot(accuracy,main="Model Accuracy", xlab="k",ylab="Accuracy")
abline(h =max(accuracy),col="red")
abline(h =min(accuracy),col="red")
```


### 5 Random Forest Algorithm for the Prediction

The 10-fold cross validation shows that random forest algorithm performs veryy well with accuracy range of 0.9964 to 0.9969. So, we use as the model for generating prediction for testing data.

```{r prediction}
rf <- randomForest(classe ~ ., data=pml.training)
varImpPlot(rf)
prediction <- predict(rf, pml.testing)
prediction 
```

